Errata

### 2025-04-05 tex/Manuscript-1.tex
Refactor code structure for improved readability and maintainability

Bulk density (BD) showed a significant difference between the study sites, with higher values in the two dryer sites, with \SI{1.5}{\gram\per\cubic\centi\meter} in PA, and \SI{1.6}{\gram\per\cubic\centi\meter} in SG, and a decrease in the more humid sites, with \SI{1.2}{\gram\per\cubic\centi\meter} in LC and \SI{0.6}{\gram\per\cubic\centi\meter} in NA. Biocrust showed no changes for BD in SG and NA, but in PA BSC+ resulted in a reduction of the BD of 18.2\%, while in LC BSC+ showed an increase in BD of 21.9\% (Figure \ref{fig:soil_properties_panel}).

Total carbon (C\textsubscript{t}) content was directly proportional to the humidity along the climatic gradient (Figure \ref{fig:soil_properties_panel}), with values of 1.1\% in PA, 0.8\% in SG, 5.0\% in LC, and 12.5\% in NA, but with a significant decrease when the BSC+ is present, from 5.6\% to 4.2\% in average. Soil inorganic carbon (SIC) was significantly higher in PA, with an average of 0.8\% of the mass of the soil, while in SG, LC, and NA, it was not found. When looking at the SIC along the different sites and under biocrust in combination, the BSC+ results in a reduction of 70.1\% in the SIC for PA, while in SG, LC and NA was no change. Soil organic carbon (SOC) showed a similar pattern as C\textsubscript{t}, with a reduction in PA to 0.3\% and the same 0.8\% in SG, 5.0\% in LC, and 12.5\%. Total nitrogen (N\textsubscript{t}) content was directly proportional to the humidity along the climatic gradient, with values of 0.04\% for PA, 0.07\% for SG, 0.28\% for LC and 0.51\% for (Figure \ref{fig:soil_properties_panel}). BSC+ showed a reduction of 30.3\% in LC and 22.3\% in NA, while the N\textsubscript{t} content remained stable in PA and SG under the influence of BSC. The relation between C\textsubscript{t} and N\textsubscript{t}, expressed as C/N, showed significantly different values of 33.9 in PA, 12.3 in SG, 16.7 in LC, and 24.5 in NA on average; and with a significant decrease from 27.0 in BSC- to 16.7 in BSC+ (Figure \ref{fig:soil_properties_panel}). It is important to note that although PA and NA present the highest values, the condition changes diametrically when observed together with the BSC+ treatment, with a large dispersion in PA and stable values in NA.

The distribution of the soil particle size classes did not show clear patterns along the climatic gradient, with PA deviating from it in all cases. Despite this, the observed values were significant, with clay values of 9.6\% for PA, 7.3\% in SG, 10.4\% in LC, and 24.6\% in NA, while when looking at the interaction between site and biocrust, there is an increase of 143.4\% in clay content for the BSC+ condition, while for SG, LC and NA was no difference; silt with 28.9\% in PA, 18.7\% in SG, 20.0\% in LC and 21.9\% in NA; and sand with 61.5\% in PA, 73.9\% in SG, 69.6\% in LC and 53.5\% in NA. When biocrusts were present, a significant decrease from 13.6\% to 12.6\% in clays and an increase from 21.2\% to 23.6\% in silt was observed, with higher dispersion for the arid site (PA) (Figure \ref{fig:soil_properties_panel}).

@@

Soil aggregate stability was evaluated through different indexes to integrate the different sizes and sieving conditions in a summary value. The difference in mean weight diameter of the aggregates ($\Delta$MWD) showed no significance in any of the conditions (Figure \ref{fig:M1-F5}). However, there was a significant difference in the difference of the geometric mean diameter of the aggregates ($\Delta$GMD) for the study sites, with values of \SI{1.86}{\milli\meter} for PA, \SI{1.2}{\milli\meter} for SG, \SI{1.4}{\milli\meter} for LC, in comparison with the more stable condition of \SI{0.83}{\milli\meter} for NA. The water-stable aggregates ratio was significant for the study sites, showing differences between NA with 81.1\% and the other study sites (57.7\% PA, 66.2\% SG, and 73.4\% LC). The ratio of soil macroaggregates of a diameter less than \SI{2.0}{\milli\meter} (R\textsubscript{$<$\SI{1.4}{\milli\meter}}) presents differences in the study sites and for biocrust treatments. SG showed a value of R\textsubscript{<\SI{2}{\milli\meter}} of 79.6\% and NA of 41.9\%, which was different from PA and LC, with 58.6\% and 62.4\%, respectively. For biocrust treatments, R\textsubscript{$<$\SI{1.4}{\milli\meter}} changed from 63.7\% to 57.5\% with biocrusts, indicating a biocrust-induced decrease in the proportion for this fraction. Finally, according to the analyzed indexes, NA showed the most stable conditions, and alternating PA and SG showed the most unstable conditions.

@@

Soil aggregate stability showed clear differences along the climatic gradient. However, when considering the effect of biocrusts, differences were limited to the smallest aggregate size class (R\textsubscript{$<$\SI{1.4}{\milli\meter}) referring to changes in microaggregates size distribution as described by \citet{Totsche2018}. Further, a difference for wet sieved aggregates with and without biocrusts between 9.5 and \SI{30.0}{\milli\meter} point to biocrust-induced stabilization of soil aggregates.

@@

This study aimed to investigate how and to what extent biocrusts stabilize soil aggregates along an arid to humid climatic gradient in Chile. We show that biocrusts play a role in soil aggregate stability along the climatic gradient by increasing the stability for wet aggregates from \SIrange[range-units=single]{9.5}{30.0}{\milli\meter} and $>$\SI{2}{\milli\meter}. Bicorusts also modify other soil properties such as C\textsubscript{t}, N\textsubscript{t}, C/N, clay, and sand, indicating an initial accumulation of organic matter that then give place to aggregation processes. The biocrusts effect on aggregates stability was lower under humid climate conditions, which indicates a transition in the main biotic agents driving the aggregation of the soil surface, moving from biocrust communities in arid regions to vascular plants in humid conditions.
Finally, we conclude that the biocrusts in our study area showed to be a valuable agent in stabilizing the upper topsoil layer, but for a narrow spectrum of conditions and mostly under arid conditions. Therefore, the effect could be considered a transition in ecological succession toward a stable ecosystem. In this process, biocrusts improve conditions for other more demanding species, such as vascular plants, initially improving the availability of carbon and nitrogen in the soil.

@@

\begin{table}[htbp] % [htbp] placement specifier (Here, Top, Bottom, Page)
\centering % Center the table on the page
\caption{Appendix A. Significant factors for response variables based on generalized linear models. Models with significant interaction also include the predictors as simple parameters based on marginality principle. \\ % Line break in caption
\textit{(p-value $\le$ 0: "***"; p-value $\le$ 0.001: "**"; p-value $\le$ 0.01: "*")}} % Significance key (using <= for clarity)
\label{tab:appendix_a} % Label for cross-referencing
\begin{tabular}{@{}lllll@{}} % Left-aligned columns, @{} removes extra space at edges
\toprule
Dependent variable & Distribution & \multicolumn{3}{c}{Significance for independent variable} \\
\cmidrule(lr){3-5} % Rule under the multicolumn header only (trimmed left and right)
 & (link-function for GLM) & Site & Biocrust & Site $\times$ Biocrust \\ % Use $\times$ for multiplication
\midrule
Clay	& Gaussian	& ***	& *	& \\
Silt	& Inverse Gaussian	& ***	& *	& \\
Sand	& Gaussian	& ***	& 	& \\
Fine Silt	& Tweedie	& ***	& 	& * \\
Medium silt	& Tweedie	& ***	& ***	& * \\
Coarse silt	& Tweedie	& ***	& 	&  \\
Very fine sand	& Gaussian	& ***	& 	&  \\
Fine sand	& Tweedie	& ***	& 	&  \\
Medium sand	& Gaussian	& ***	& 	&  \\
Coarse sand	& Tweedie	& ***	& 	&  \\
pH	& Gaussian	& ***	& **	&  \\
EC	& Tweedie	& ***	& 	&  \\
BD	& Gaussian	& ***	& 	& ** \\
N\textsubscript{t}	& Gamma	& ***	& 	& ** \\ % For Nt subscript
C\textsubscript{t}	& Tweedie	& ***	& **	&  \\ % For Ct subscript
C/N	& Tweedie	& ***	& **	&  \\
19.0-30.0 mm Dry	& Tweedie	& ***	& 	&  \\
9.5-19.0 mm Dry	& Tweedie	& ***	& 	& ** \\
6.7-9.5 mm Dry	& Tweedie	& ***	& 	& ** \\
4.7-6.7 mm Dry	& Tweedie	& ***	& 	& * \\
3.4-4.7 mm Dry	& Tweedie	& ***	& 	&  \\
2.0-3.4 mm Dry	& Tweedie	& ***	& 	& * \\
\textless2.0 mm Dry	& Tweedie	& ***	& 	& *** \\ % Use \textless for less than symbol
19.0-30.0 mm Wet	& Tweedie	& ***	& *	& ** \\
9.5-19.0 mm Wet	& Tweedie	& *	& **	& * \\
6.7-9.5 mm Wet	& Tweedie	& ***	& 	&  \\
4.7-6.7 mm Wet	& Tweedie	& ***	& 	&  \\
3.4-4.7 mm Wet	& Tweedie	& ***	& 	&  \\
2.0-3.4 mm Wet	& Tweedie	& ***	& 	&  \\
\textless2.0 mm Wet	& Tweedie	& ***	& *	&  \\ % Use \textless for less than symbol
$\Delta$MWD	& Gaussian	& 	& 	&  \\ % Use \Delta for Greek Delta
$\Delta$GMD	& Tweedie	& ***	& 	&  \\ % Use \Delta for Greek Delta
WSAR	& Gaussian	& 	& 	&  \\
R\textless2 mm	& Gaussian	& ***	& *	&  \\ % Use \textless for less than symbol
\bottomrule
\end{tabular}
\end{table}

>>>>

Bulk density (BD) showed a significant difference between the study sites, with higher values in the two dryer sites, with \SI{1.5}{\gram\per\cubic\centi\meter} in PA, and \SI{1.6}{\gram\per\cubic\centi\meter} in SG, and a decrease in the more humid sites, with \SI{1.2}{\gram\per\cubic\centi\meter} in LC and \SI{0.6}{\gram\per\cubic\centi\meter} in NA. Biocrust showed no changes for BD in SG and NA, but in PA BSC+ resulted in a reduction of the BD of 18.2\%, while in LC BSC+ showed an increase in BD of 21.9\% (Figure \ref{fig:M1-F2-soil_properties_panel}).

Total carbon (C\textsubscript{t}) content was directly proportional to the humidity along the climatic gradient (Figure \ref{fig:M1-F2-soil_properties_panel}), with values of 1.1\% in PA, 0.8\% in SG, 5.0\% in LC, and 12.5\% in NA, but with a significant decrease when the BSC+ is present, from 5.6\% to 4.2\% in average. Soil inorganic carbon (SIC) was significantly higher in PA, with an average of 0.8\% of the mass of the soil, while in SG, LC, and NA, it was not found. When looking at the SIC along the different sites and under biocrust in combination, the BSC+ results in a reduction of 70.1\% in the SIC for PA, while in SG, LC and NA was no change. Soil organic carbon (SOC) showed a similar pattern as C\textsubscript{t}, with a reduction in PA to 0.3\% and the same 0.8\% in SG, 5.0\% in LC, and 12.5\%. Total nitrogen (N\textsubscript{t}) content was directly proportional to the humidity along the climatic gradient, with values of 0.04\% for PA, 0.07\% for SG, 0.28\% for LC and 0.51\% for (Figure \ref{fig:M1-F2-soil_properties_panel}). BSC+ showed a reduction of 30.3\% in LC and 22.3\% in NA, while the N\textsubscript{t} content remained stable in PA and SG under the influence of BSC. The relation between C\textsubscript{t} and N\textsubscript{t}, expressed as C/N, showed significantly different values of 33.9 in PA, 12.3 in SG, 16.7 in LC, and 24.5 in NA on average; and with a significant decrease from 27.0 in BSC- to 16.7 in BSC+ (Figure \ref{fig:M1-F2-soil_properties_panel}). It is important to note that although PA and NA present the highest values, the condition changes diametrically when observed together with the BSC+ treatment, with a large dispersion in PA and stable values in NA.

The distribution of the soil particle size classes did not show clear patterns along the climatic gradient, with PA deviating from it in all cases. Despite this, the observed values were significant, with clay values of 9.6\% for PA, 7.3\% in SG, 10.4\% in LC, and 24.6\% in NA, while when looking at the interaction between site and biocrust, there is an increase of 143.4\% in clay content for the BSC+ condition, while for SG, LC and NA was no difference; silt with 28.9\% in PA, 18.7\% in SG, 20.0\% in LC and 21.9\% in NA; and sand with 61.5\% in PA, 73.9\% in SG, 69.6\% in LC and 53.5\% in NA. When biocrusts were present, a significant decrease from 13.6\% to 12.6\% in clays and an increase from 21.2\% to 23.6\% in silt was observed, with higher dispersion for the arid site (PA) (Figure \ref{fig:M1-F2-soil_properties_panel}).

@@

Soil aggregate stability was evaluated through different indexes to integrate the different sizes and sieving conditions in a summary value. The difference in mean weight diameter of the aggregates ($\Delta$MWD) showed no significance in any of the conditions (Figure \ref{fig:M1-F5}). However, there was a significant difference in the difference of the geometric mean diameter of the aggregates ($\Delta$GMD) for the study sites, with values of \SI{1.86}{\milli\meter} for PA, \SI{1.2}{\milli\meter} for SG, \SI{1.4}{\milli\meter} for LC, in comparison with the more stable condition of \SI{0.83}{\milli\meter} for NA. The water-stable aggregates ratio was significant for the study sites, showing differences between NA with 81.1\% and the other study sites (57.7\% PA, 66.2\% SG, and 73.4\% LC). The ratio of soil macroaggregates of a diameter less than \SI{2.0}{\milli\meter} (R\textsubscript{$<$\SI{0.0}{\milli\meter}}) presents differences in the study sites and for biocrust treatments. SG showed a value of R\textsubscript{<\SI{2}{\milli\meter}} of 79.6\% and NA of 41.9\%, which was different from PA and LC, with 58.6\% and 62.4\%, respectively. For biocrust treatments, R\textsubscript{$<$\SI{2.0}{\milli\meter}} changed from 63.7\% to 57.5\% with biocrusts, indicating a biocrust-induced decrease in the proportion for this fraction. Finally, according to the analyzed indexes, NA showed the most stable conditions, and alternating PA and SG showed the most unstable conditions.

@@

Soil aggregate stability showed clear differences along the climatic gradient. However, when considering the effect of biocrusts, differences were limited to the smallest aggregate size class (R\textsubscript{$<$\SI{2.0}{\milli\meter}}) referring to changes in microaggregates size distribution as described by \citet{Totsche2018}. Further, a difference for wet sieved aggregates with and without biocrusts between 9.5 and \SI{30.0}{\milli\meter} point to biocrust-induced stabilization of soil aggregates.

@@

This study aimed to investigate how and to what extent biocrusts stabilize soil aggregates along an arid to humid climatic gradient in Chile. We show that biocrusts play a role in soil aggregate stability along the climatic gradient by increasing the stability for wet aggregates from \SIrange[range-units=single]{9.5}{30.0}{\milli\meter} and $>$\SI{2}{\milli\meter}. Bicorusts also modify other soil properties such as C\textsubscript{t}, N\textsubscript{t}, C/N, clay, and sand, indicating an initial accumulation of organic matter that then give place to aggregation processes. The biocrusts effect on aggregates stability was lower under humid climate conditions, which indicates a transition in the main biotic agents driving the aggregation of the soil surface, moving from biocrust communities in arid regions to vascular plants in humid conditions.

Finally, we conclude that the biocrusts in our study area showed to be a valuable agent in stabilizing the upper topsoil layer, but for a narrow spectrum of conditions and mostly under arid conditions. Therefore, the effect could be considered a transition in ecological succession toward a stable ecosystem. In this process, biocrusts improve conditions for other more demanding species, such as vascular plants, initially improving the availability of carbon and nitrogen in the soil.

@@

\begin{table}[htbp] % [htbp] placement specifier (Here, Top, Bottom, Page)
\centering % Center the table on the page
\caption{Appendix A. Significant factors for response variables based on generalized linear models. Models with significant interaction also include the predictors as simple parameters based on marginality principle. \\ % Line break in caption
\textit{(p-value $\le$ 0: "***"; p-value $\le$ 0.001: "**"; p-value $\le$ 0.01: "*")}} % Significance key (using <= for clarity)
\label{tab:appendix_a} % Label for cross-referencing
\begin{tabular}{@{}lllll@{}} % Left-aligned columns, @{} removes extra space at edges
\toprule
Dependent variable & Distribution & \multicolumn{3}{c}{Significance for independent variable} \\
\cmidrule(lr){3-5} % Rule under the multicolumn header only (trimmed left and right)
 & (link-function for GLM) & Site & Biocrust & Site $\times$ Biocrust \\ % Use $\times$ for multiplication
\midrule
Clay	& Gaussian	& ***	& *	& \\
Silt	& Inverse Gaussian	& ***	& *	& \\
Sand	& Gaussian	& ***	& 	& \\
Fine Silt	& Tweedie	& ***	& 	& * \\
Medium silt	& Tweedie	& ***	& ***	& * \\
Coarse silt	& Tweedie	& ***	& 	&  \\
Very fine sand	& Gaussian	& ***	& 	&  \\
Fine sand	& Tweedie	& ***	& 	&  \\
Medium sand	& Gaussian	& ***	& 	&  \\
Coarse sand	& Tweedie	& ***	& 	&  \\
pH	& Gaussian	& ***	& **	&  \\
EC	& Tweedie	& ***	& 	&  \\
BD	& Gaussian	& ***	& 	& ** \\
N\textsubscript{t}	& Gamma	& ***	& 	& ** \\ % For Nt subscript
C\textsubscript{t}	& Tweedie	& ***	& **	&  \\ % For Ct subscript
C/N	& Tweedie	& ***	& **	&  \\
19.0-30.0 mm Dry	& Tweedie	& ***	& 	&  \\
9.5-19.0 mm Dry	& Tweedie	& ***	& 	& ** \\
6.7-9.5 mm Dry	& Tweedie	& ***	& 	& ** \\
4.7-6.7 mm Dry	& Tweedie	& ***	& 	& * \\
3.4-4.7 mm Dry	& Tweedie	& ***	& 	&  \\
2.0-3.4 mm Dry	& Tweedie	& ***	& 	& * \\
\textless2.0 mm Dry	& Tweedie	& ***	& 	& *** \\ % Use \textless for less than symbol
19.0-30.0 mm Wet	& Tweedie	& ***	& *	& ** \\
9.5-19.0 mm Wet	& Tweedie	& *	& **	& * \\
6.7-9.5 mm Wet	& Tweedie	& ***	& 	&  \\
4.7-6.7 mm Wet	& Tweedie	& ***	& 	&  \\
3.4-4.7 mm Wet	& Tweedie	& ***	& 	&  \\
2.0-3.4 mm Wet	& Tweedie	& ***	& 	&  \\
\textless2.0 mm Wet	& Tweedie	& ***	& *	&  \\ % Use \textless for less than symbol
$\Delta$MWD	& Gaussian	& 	& 	&  \\ % Use \Delta for Greek Delta
$\Delta$GMD	& Tweedie	& ***	& 	&  \\ % Use \Delta for Greek Delta
WSAR	& Gaussian	& 	& 	&  \\
R\textsubscript{2 mm}	& Gaussian	& ***	& *	&  \\ % Use \textless for less than symbol
\bottomrule
\end{tabular}
\end{table}
